---
title: "Initiation_Sequence_Context_Analysis"
author: "Xiaoting Li"
date: "December 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data   
   
CDS sequences: 100bp around start sites  
uORF sequences: 100bp around start sites  
```{r message=FALSE, echo=FALSE}
library(ggseqlogo)
library(ggplot2)
library(rtracklayer)
library(Biostrings)

Dir <- "~/Desktop/uORF/"
Files.Dir <- "~/Desktop/uORF/FINAL_Bed_and_FASTA_Files/"

uorf.extended <- import(paste(Files.Dir, "uORFs_100bpwindows_aroundstart.fasta", sep=""), format="fasta")
cds.extended <- import(paste(Files.Dir, "CDS_100bpwindows_aroundstart.fasta", sep=""), format="fasta")
```
   
## uORF start codon usage   
     
```{r message=FALSE, warning=FALSE, echo=FALSE}
uorf.start <- sapply(1:length(uorf.extended), function(i)strsplit(names(uorf.extended)[i], split="_")[[1]][5])
uorf.start <- as.numeric(uorf.start)

uorf.start.codon <- as.character(subseq(uorf.extended, uorf.start+1, uorf.start+3))
table(uorf.start.codon) 

ggplot(data.frame(uorf.start.codon), aes(uorf.start.codon)) +
  geom_histogram(stat="count") +
  theme_light()
```
    
Probability     
```{r echo=FALSE}
ggseqlogo(uorf.start.codon, method = "probability")
``` 
    
Information content       
```{r echo=FALSE}
ggseqlogo(uorf.start.codon)
```
    

## Sequence context around  initiation start codon   
     
###CDS initiation sequence context   
    
Kozak consensus sequence:
https://en.wikipedia.org/wiki/Kozak_consensus_sequence   
    
```{r echo=FALSE}
cds.start <- sapply(1:length(cds.extended), function(i)strsplit(names(cds.extended)[i], split="_")[[1]][4])
cds.start <- as.numeric(cds.start)

cds.ini.minus <- as.character(subseq(cds.extended, cds.start-9, cds.start))
cds.ini.plus <- as.character(subseq(cds.extended, cds.start+4, cds.start+13))
cds.ini <- paste(cds.ini.minus, "   ", cds.ini.plus, sep="")
ggseqlogo(cds.ini, method = 'bits') +
  annotate("text", x=11:13, y=0.02, label =c("A", "T", "G"), size=5) +
  labs(title = "CDS") +
  theme(plot.title = element_text(hjust = 0.5))
```
   
###uORF initiation sequence context

All uORFs
```{r echo=FALSE}
uorf.ini.minus <- as.character(subseq(uorf.extended, uorf.start-9, uorf.start))
uorf.ini.plus <- as.character(subseq(uorf.extended, uorf.start+4, uorf.start+13))
uorf.ini <- paste(uorf.ini.minus, "   ", uorf.ini.plus, sep="")
ggseqlogo(uorf.ini, method = 'bits') +
  annotate("text", x=12, y=0.02, label = "start", size=4) +
  annotate("text", x=12, y=0.01, label = "codon", size=4) +
  labs(title = "all uORFs") +
  theme(plot.title = element_text(hjust = 0.5))
```
    
AUG start codon uORFs vs. Non-AUG start codon uORFs 
```{r echo=FALSE}
uorf.ini.list1 <- list(ATG = uorf.ini[which(uorf.start.codon == "ATG")], Non.ATG = uorf.ini[which(uorf.start.codon != "ATG")])

ggseqlogo(uorf.ini.list1, ncol = 1)
```
     
uORFs grouped by start codon     
```{r echo=FALSE}
# get rid of AAG and AGG: few uORFs
codons <- unique(uorf.start.codon)
codons <- codons[1:8]
uorf.ini.list <- lapply(1:length(codons), function(i) uorf.ini[uorf.start.codon == codons[i]])
names(uorf.ini.list) <- codons

ggseqlogo(uorf.ini.list, ncol = 2)
```

